---
title: "iNEXT"
output: html_document
date: "2024-09-03"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE)
opts_knit$set(root.dir='../')                     # definining working directory; or normalizePath('../')
options(tibble.print_max = 200, tibble.print_min = 20)
```

# load libaries
```{r}
#Install/load pacman
if(!require(pacman)){install.packages("pacman");require(pacman)}
#Install packages
p_load("tidyverse","vegan","reshape2","magrittr","caret","ggplot2",
       "iNEXT","ggpubr","ggpmisc")
```


# load rawdata
```{r}
load("data/workingdata.RData")
```

# data preparation for iNEXT
A001 = b	Bindalm
A002 = e	Engertalm
A003 = h	Halsalm
A004 = k	Königsbachalm1
A005 = a	Königsbachalm2
A006 = m	Königsbergalm
A007 = m	Mittereisalm
A008 = s	Saletalm

```{r}
df = vegTreatment%>% group_by(treatment,pasture,plot,species)%>% dplyr::select(treatment,pasture,plot_id,species)
df = df[!duplicated(df), ]

vegNspeciesTreatment = df %>% 
  dplyr::group_by(treatment,pasture,species) %>% 
  dplyr::summarise(n = n())

vegNspeciesTreatment = data.frame(vegNspeciesTreatment)
```

### control
#### complete dataset
```{r}
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(treatment =="control") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="total",
                                         species="",
                                         n=72),
                              vegNspeciesTreatmentN)

control = vegNspeciesTreatmentN$n
names(control) =  vegNspeciesTreatmentN$species
```

#### per grassland pasture
```{r}
# define data for control; 9 plots in total per pasture
# A001 - Bindalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>% dplyr::filter(treatment =="control",
                                                               pasture=="Bindalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Bindalm",species="",n=9),vegNspeciesTreatmentN)

A001_C = vegNspeciesTreatmentN$n
names(A001_C) =  vegNspeciesTreatmentN$species

# A002 - Engertalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>% dplyr::filter(treatment =="control",
                                                               pasture=="Engertalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Engertalm",species="",n=9),vegNspeciesTreatmentN)

A002_C = vegNspeciesTreatmentN$n
names(A002_C) =  vegNspeciesTreatmentN$species


# A003 - Halsalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>% dplyr::filter(treatment =="control",
                                                               pasture=="Halsalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Halsalm",species="",n=9),vegNspeciesTreatmentN)

A003_C = vegNspeciesTreatmentN$n
names(A003_C) =  vegNspeciesTreatmentN$species


# A004 - Königsbachalm1
vegNspeciesTreatmentN = vegNspeciesTreatment %>% dplyr::filter(treatment =="control",
                                                               pasture=="Königsbachalm 1") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Königsbachalm1",species="",n=9),vegNspeciesTreatmentN)

A004_C = vegNspeciesTreatmentN$n
names(A004_C) =  vegNspeciesTreatmentN$species

# A005 - Königsbachalm2
vegNspeciesTreatmentN = vegNspeciesTreatment %>% dplyr::filter(treatment =="control",
                                                               pasture=="Königsbachalm 2") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Königsbachalm2",species="",n=9),vegNspeciesTreatmentN)

A005_C = vegNspeciesTreatmentN$n
names(A005_C) =  vegNspeciesTreatmentN$species



# A006 - Königsbergalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>% dplyr::filter(treatment =="control",
                                                               pasture=="Königsbergalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Königsbergalm",species="",n=9),vegNspeciesTreatmentN)

A006_C = vegNspeciesTreatmentN$n
names(A006_C) =  vegNspeciesTreatmentN$species

# A007 - Mittereisalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>% dplyr::filter(treatment =="control",
                                                               pasture=="Mittereisalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Mittereisalm",species="",n=9),vegNspeciesTreatmentN)

A007_C = vegNspeciesTreatmentN$n
names(A007_C) =  vegNspeciesTreatmentN$species

# A008 - Saletalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>% dplyr::filter(treatment =="control",
                                                               pasture=="Saletalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Mittereisalm",species="",n=9),vegNspeciesTreatmentN)

A008_C = vegNspeciesTreatmentN$n
names(A008_C) =  vegNspeciesTreatmentN$species
```



## encroachment
#### create complete dateset
```{r}
vegNspeciesTreatmentN = vegNspeciesTreatment %>% 
  dplyr::filter(treatment =="encroachment") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="total",
                species="",
                n=72),
                vegNspeciesTreatmentN)

encroachment = vegNspeciesTreatmentN$n
names(encroachment) =  vegNspeciesTreatmentN$species
```


#### create datasets per pasture
```{r}
# define data for encroachment; 9 plots in total per pasture
# A001 - Bindalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>% 
  dplyr::filter(treatment =="encroachment",
                 pasture=="Bindalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Bindalm",
                                         species="",n=9),
                              vegNspeciesTreatmentN)

A001_V = vegNspeciesTreatmentN$n
names(A001_V) =  vegNspeciesTreatmentN$species

# A002 - Engertalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(treatment =="encroachment",
         pasture=="Engertalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Engertalm",species="",n=9),
                              vegNspeciesTreatmentN)

A002_V = vegNspeciesTreatmentN$n
names(A002_V) =  vegNspeciesTreatmentN$species


# A003 - Halsalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(treatment =="encroachment",
               pasture=="Halsalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Halsalm",
                                         species="",n=9),
                              vegNspeciesTreatmentN)

A003_V = vegNspeciesTreatmentN$n
names(A003_V) =  vegNspeciesTreatmentN$species

# A004 - Königsbachalm1
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(treatment =="encroachment",
                pasture=="Königsbachalm 1") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Königsbachalm1",
                                         species="",n=9),
                              vegNspeciesTreatmentN)

A004_V = vegNspeciesTreatmentN$n
names(A004_V) =  vegNspeciesTreatmentN$species

# A005 - Königsbachalm2
vegNspeciesTreatmentN = vegNspeciesTreatment %>% 
  dplyr::filter(treatment =="encroachment",
                pasture=="Königsbachalm 2") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Königsbachalm2",
                                         species="",n=9),
                              vegNspeciesTreatmentN)

A005_V = vegNspeciesTreatmentN$n
names(A005_V) =  vegNspeciesTreatmentN$species

# A006 - Königsbergalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(treatment =="encroachment",
                                                               pasture=="Königsbergalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Königsbergalm",
                                         species="",n=9),
                              vegNspeciesTreatmentN)

A006_V = vegNspeciesTreatmentN$n
names(A006_V) =  vegNspeciesTreatmentN$species

# A007 - Mittereisalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(treatment =="encroachment",
              pasture=="Mittereisalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Mittereisalm",
                                         species="",n=9),
                              vegNspeciesTreatmentN)

A007_V = vegNspeciesTreatmentN$n
names(A007_V) =  vegNspeciesTreatmentN$species

# A008 - Saletalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(treatment =="encroachment",
                  pasture=="Saletalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Mittereisalm",
                                         species="",n=9),
                              vegNspeciesTreatmentN)

A008_V = vegNspeciesTreatmentN$n
names(A008_V) =  vegNspeciesTreatmentN$species
```




```{r}
# prepare for iNEXT
vegInput = list(
  control = control, 
  encroachment = encroachment)


vegInput_A001 = list(
  control = A001_C, 
  encroachment = A001_V)

vegInput_A002 = list(
  control = A002_C, 
  encroachment = A002_V)

vegInput_A003 = list(
  control = A003_C, 
  encroachment = A003_V)

vegInput_A004 = list(
  control = A004_C, 
  encroachment = A004_V)

vegInput_A005 = list(
  control = A005_C, 
  encroachment = A005_V)

vegInput_A006 = list(
  control = A006_C, 
  encroachment = A006_V)

vegInput_A007 = list(
  control = A007_C, 
  encroachment = A007_V)

vegInput_A008 = list(
  control = A008_C, 
  encroachment = A008_V)

```


# iNEXT
## All pastures together
```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2024)

t <- seq(1, 72, by=1)
out.inc <- iNEXT(vegInput, q=0, datatype="incidence_freq", size=t)
# Sample‐size‐based R/E curves

gt = ggiNEXT(out.inc, type=1, color.var="Assemblage") +
theme_bw(base_size = 18) +
theme(legend.position="bottom",
legend.title=element_blank(),
text=element_text(size=18),
legend.box = "vertical") +
  #https://jrgraphix.net/r/Unicode/25A0-25FF
  scale_shape_manual(values=c("\u25B3","\u25B2"))+
    scale_fill_manual(values=c(
"#FF0000",
"#0000FF"))+
  scale_colour_manual(values=c(
"#FF0000",
"#0000FF"))+
  scale_x_continuous(n.breaks = 10, 
                     limits = c(1, 72))+
  labs(title="");gt
```


## Bindalm
```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2024)
t <- seq(1, 9, by=1)
out.inc <- iNEXT(vegInput_A001, q=0, datatype="incidence_freq", size=t)
# Sample‐size‐based R/E curves

g1 = ggiNEXT(out.inc, type=1, color.var="Assemblage") +
theme_bw(base_size = 18) +
theme(legend.position="bottom",
legend.title=element_blank(),
text=element_text(size=18),
legend.box = "vertical") +
  #https://jrgraphix.net/r/Unicode/25A0-25FF
  scale_shape_manual(values=c("\u25B3","\u25B2"))+
    scale_fill_manual(values=c(
"#FF0000",
"#0000FF"))+
  scale_colour_manual(values=c(
"#FF0000",
"#0000FF"))+
  scale_x_continuous(n.breaks = 9, limits = c(1, 9))+
  labs(title="Bindalm (1100 m)")
```

## Engertalm

```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2024)
t <- seq(1, 9, by=1)
out.inc <- iNEXT(vegInput_A002, q=0, datatype="incidence_freq", size=t)
# Sample‐size‐based R/E curves

g2 = ggiNEXT(out.inc, type=1, color.var="Assemblage") +
theme_bw(base_size = 18) +
theme(legend.position="bottom",
legend.title=element_blank(),
text=element_text(size=18),
legend.box = "vertical") +
  #https://jrgraphix.net/r/Unicode/25A0-25FF
  scale_shape_manual(values=c("\u25B3","\u25B2"))+
    scale_fill_manual(values=c(
"#FF0000","#0000FF"))+
  scale_colour_manual(values=c(
"#FF0000","#0000FF"))+
  scale_x_continuous(n.breaks = 9, limits = c(1, 9))+
  labs(title="Engertalm (995 m)")
```


## Halsalm

```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2024)
t <- seq(1, 9, by=1)
out.inc <- iNEXT(vegInput_A003, q=0, datatype="incidence_freq", size=t)
# Sample‐size‐based R/E curves

g3 = ggiNEXT(out.inc, type=1, color.var="Assemblage") +
theme_bw(base_size = 18) +
theme(legend.position="bottom",
legend.title=element_blank(),
text=element_text(size=18),
legend.box = "vertical") +
  #https://jrgraphix.net/r/Unicode/25A0-25FF
  scale_shape_manual(values=c("\u25B3","\u25B2"))+
    scale_fill_manual(values=c(
"#FF0000","#0000FF"))+
  scale_colour_manual(values=c(
"#FF0000","#0000FF"))+
  scale_x_continuous(n.breaks = 9, limits = c(1, 9))+
  labs(title="Halsalm (1250 m)")
```


## 	Königsbachalm 1

```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2024)
t <- seq(1, 9, by=1)
out.inc <- iNEXT(vegInput_A004, q=0, datatype="incidence_freq", size=t)
# Sample‐size‐based R/E curves

g4 = ggiNEXT(out.inc, type=1, color.var="Assemblage") +
theme_bw(base_size = 18) +
theme(legend.position="bottom",
legend.title=element_blank(),
text=element_text(size=18),
legend.box = "vertical") +
  #https://jrgraphix.net/r/Unicode/25A0-25FF
  scale_shape_manual(values=c("\u25B3","\u25B2"))+
    scale_fill_manual(values=c(
"#FF0000","#0000FF"))+
  scale_colour_manual(values=c(
"#FF0000","#0000FF"))+
  scale_x_continuous(n.breaks = 9, limits = c(1, 9))+
  labs(title="	Königsbachalm 1 (1200 m)")
```


## Königsbachalm 2

```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2024)
t <- seq(1, 9, by=1)
out.inc <- iNEXT(vegInput_A005, q=0, datatype="incidence_freq", size=t)
# Sample‐size‐based R/E curves

g5 = ggiNEXT(out.inc, type=1, color.var="Assemblage") +
theme_bw(base_size = 18) +
theme(legend.position="bottom",
legend.title=element_blank(),
text=element_text(size=18),
legend.box = "vertical") +
  #https://jrgraphix.net/r/Unicode/25A0-25FF
  scale_shape_manual(values=c("\u25B3","\u25B2"))+
    scale_fill_manual(values=c(
"#FF0000","#0000FF"))+
  scale_colour_manual(values=c(
"#FF0000","#0000FF"))+
  scale_x_continuous(n.breaks = 9, limits = c(1, 9))+
  labs(title="Königsbachalm 2 (1270 m)")
```

## Königsbergalm
```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2024)
t <- seq(1, 9, by=1)
out.inc <- iNEXT(vegInput_A006, q=0, datatype="incidence_freq", size=t)
# Sample‐size‐based R/E curves

g6 = ggiNEXT(out.inc, type=1, color.var="Assemblage") +
theme_bw(base_size = 18) +
theme(legend.position="bottom",
legend.title=element_blank(),
text=element_text(size=18),
legend.box = "vertical") +
  #https://jrgraphix.net/r/Unicode/25A0-25FF
  scale_shape_manual(values=c("\u25B3","\u25B2"))+
    scale_fill_manual(values=c(
"#FF0000", "#0000FF"))+
  scale_colour_manual(values=c(
"#FF0000","#0000FF"))+
  scale_x_continuous(n.breaks = 9, limits = c(1, 9))+
  labs(title="Königsbergalm (1595 m)")
```

## Mittereisalm
```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2024)
t <- seq(1, 9, by=1)
out.inc <- iNEXT(vegInput_A007, q=0, datatype="incidence_freq", size=t)
# Sample‐size‐based R/E curves

g7 = ggiNEXT(out.inc, type=1, color.var="Assemblage") +
theme_bw(base_size = 18) +
theme(legend.position="bottom",
legend.title=element_blank(),
text=element_text(size=18),
legend.box = "vertical") +
  #https://jrgraphix.net/r/Unicode/25A0-25FF
  scale_shape_manual(values=c("\u25B3","\u25B2"))+
    scale_fill_manual(values=c(
"#FF0000","#0000FF"))+
  scale_colour_manual(values=c(
"#FF0000","#0000FF"))+
  scale_x_continuous(n.breaks = 9, limits = c(1, 9))+
  labs(title="Mittereisalm (1340 m)")
```

## Saletalm
```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2024)
t <- seq(1, 9, by=1)
out.inc <- iNEXT(vegInput_A008, q=0, datatype="incidence_freq", size=t)
# Sample‐size‐based R/E curves

g8 = ggiNEXT(out.inc, type=1, color.var="Assemblage") +
theme_bw(base_size = 18) +
theme(legend.position="bottom",
legend.title=element_blank(),
text=element_text(size=18),
legend.box = "vertical") +
  #https://jrgraphix.net/r/Unicode/25A0-25FF
  scale_shape_manual(values=c("\u25B3","\u25B2"))+
    scale_fill_manual(values=c(
"#FF0000","#0000FF"))+
  scale_colour_manual(values=c(
"#FF0000","#0000FF"))+
  scale_x_continuous(n.breaks = 9, limits = c(1, 9))+
  labs(title="Saletalm (680 m)")
```

A001 = b	Bindalm
A002 = e	Engertalm
A003 = h	Halsalm
A004 = k	Königsbachalm1
A005 = a	Königsbachalm2
A006 = m	Königsbergalm
A007 = m	Mittereisalm
A008 = s	Saletalm



## Combined plot
```{r}
#ggarrange(g1,g2,g3,g4,g5,g6,g7,g8, ncol = 4, nrow = 2, common.legend = TRUE)
# according elevation

ggarrange(g8,g2,g1,g4,
          g3,g5,g7,g6, ncol = 4, nrow = 2, common.legend = TRUE,
legend="bottom")

```

# No distinction between control and encroachment
## Create datasets per pasture
```{r}
# define data for encroachment; 9 plots in total per pasture
# A001 - Bindalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>% 
  dplyr::filter(pasture=="Bindalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Bindalm",
                                         species="",n=18),
                              vegNspeciesTreatmentN)

b = vegNspeciesTreatmentN$n
names(b) =  vegNspeciesTreatmentN$species

# A002 - Engertalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(pasture=="Engertalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Engertalm",species="",n=18),
                              vegNspeciesTreatmentN)

e = vegNspeciesTreatmentN$n
names(e) =  vegNspeciesTreatmentN$species


# A003 - Halsalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(pasture=="Halsalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Halsalm",
                                         species="",n=18),
                              vegNspeciesTreatmentN)

h = vegNspeciesTreatmentN$n
names(h) =  vegNspeciesTreatmentN$species

# A004 - Königsbachalm1
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(pasture=="Königsbachalm 1") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Königsbachalm1",
                                         species="",n=18),
                              vegNspeciesTreatmentN)

k = vegNspeciesTreatmentN$n
names(k) =  vegNspeciesTreatmentN$species

# A005 - Königsbachalm2
vegNspeciesTreatmentN = vegNspeciesTreatment %>% 
  dplyr::filter(pasture=="Königsbachalm 2") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Königsbachalm2",
                                         species="",n=18),
                              vegNspeciesTreatmentN)

a = vegNspeciesTreatmentN$n
names(a) =  vegNspeciesTreatmentN$species

# A006 - Königsbergalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(pasture=="Königsbergalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Königsbergalm",
                                         species="",n=18),
                              vegNspeciesTreatmentN)

g = vegNspeciesTreatmentN$n
names(g) =  vegNspeciesTreatmentN$species

# A007 - Mittereisalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(pasture=="Mittereisalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Mittereisalm",
                                         species="",n=18),
                              vegNspeciesTreatmentN)

m = vegNspeciesTreatmentN$n
names(m) =  vegNspeciesTreatmentN$species

# A008 - Saletalm
vegNspeciesTreatmentN = vegNspeciesTreatment %>%
  dplyr::filter(pasture=="Saletalm") %>%
  dplyr::select(pasture,species,n)
vegNspeciesTreatmentN = rbind(data.frame(pasture="Mittereisalm",
                                         species="",n=18),
                              vegNspeciesTreatmentN)

s = vegNspeciesTreatmentN$n
names(s) =  vegNspeciesTreatmentN$species


vegInput = list(
e = e,
b = b,
s = s,
k = k,
h = h,
m = m,
a = a,
g = g)
```


## iNEXT
```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2024)
t <- seq(1, 13, by=1)
out.inc <- iNEXT(vegInput, q=0, datatype="incidence_freq", size=t)
# Sample‐size‐based R/E curves
```


## Plot iNEXT output
```{r}
g = ggiNEXT(out.inc, type=1, color.var="Assemblage") +
theme_bw(base_size = 18) +
theme(legend.position="bottom",
legend.title=element_blank(),
text=element_text(size=18),
legend.box = "vertical") +
  #https://jrgraphix.net/r/Unicode/25A0-25FF
  scale_shape_manual(values=c("\u25B3","\u25B2","\u25BD","\u25BC","\u25c7","\u25c6","\u25CB","\u25CF"))+
    scale_fill_manual(values=c(
"#FF0000","#00FF00","#00FFFF","#FF00FF",
"#0000FF","#FF8000","#800000","#808000"))+
  scale_colour_manual(values=c(
"#FF0000","#00FF00","#00FFFF","#FF00FF",
"#0000FF","#FF8000","#800000","#808000"))

g
```



```{r}
#order BBG according iNEXT
arranged = out.inc$DataInfo %>% dplyr::arrange(desc(S.obs))
arrangedSites = arranged %>% dplyr::select(Assemblage,S.obs) %>%
  dplyr::rename(site_id=Assemblage)
```


# Create Final table

```{r}
df1 = arrangedSites

df2 = out.inc$iNextEst$size_based %>% 
  dplyr::filter(t==1) %>%
  dplyr::rename(site_id=Assemblage) %>%
  dplyr::select(site_id,qD)

df_list <- list(df1, df2)      

#merge all data frames together
final = Reduce(function(x, y) merge(x, y, all.x=TRUE), df_list)

final %<>% dplyr::rename(gammaPlot = S.obs,
                        alphaPlot = qD) %>% 
  dplyr::select(site_id, alphaPlot,gammaPlot)
```


## Calculate slope

```{r}
# not correct
# https://rpubs.com/brouwern/iNEXTvVEGAN
# raremax <- min(rowSums(vegInc))
# dat = data.frame(vegan::rareslope(vegInc,sample=raremax))
# vegSlopes = data.frame(rownames(dat),dat[,1])
# colnames(vegSlopes)=c("site_id","slope")


final$ln_alpha = log(final$alphaPlot)
final$ln_gamma = log(final$gammaPlot)

final$slope = (final$ln_gamma-final$ln_alpha)/log(18)
```

## Calculate diversity estimators
1st order jackknife: It uses the frequency of uniques to estimate the number of undetected species; 
see Burnham and Overton (1978).

2nd order jackknife: It uses the frequencies of uniques and duplicates to estimate the number of undetected species; 
see Burnham and Overton (1978).

Chao2 (Chao, 1987): This approach uses the frequencies of uniques and duplicates to estimate the number of undetected species; see Chao (1987) or Eq. (11a) in Chao and Chiu (2016b).



### create table
```{r}
df = veg %>% dplyr::filter(layer=="hl")
rownames(df) = df[,1]
df = df[,-c(1,2)] %>% as.data.frame() 

e = df %>% select(starts_with("e"))
b = df %>% select(starts_with("b"))
s = df %>% select(starts_with("s"))
k = df %>% select(starts_with("k"))
h = df %>% select(starts_with("h"))
m = df %>% select(starts_with("m"))
a = df %>% select(starts_with("a"))
g = df %>% select(starts_with("g"))

vegList = list(
e = e,
b = b,
s = s,
k = k,
h = h,
m = m,
a = a,
g = g)
```


```{r}
# final$jackknife_1order = unlist(lapply(veg,fossil::jack1))
# final$jackknife_2order = unlist(lapply(veg,fossil::jack2))
# final$Chao2 = unlist(lapply(veg,fossil::chao2))


chaoEstimates = lapply(vegList,function(x) SpadeR::ChaoSpecies(x,"incidence_raw",k=10,conf=0.95))



df = data.frame(rbind(
chaoEstimates$e$Species_table,
chaoEstimates$b$Species_table,
chaoEstimates$s$Species_table,
chaoEstimates$k$Species_table,
chaoEstimates$h$Species_table,
chaoEstimates$m$Species_table,
chaoEstimates$a$Species_table,
chaoEstimates$g$Species_table))

df$site_id = rep(c("e","b","s","k","h","m","a","g"),each=8)
df$type = rownames(df)



# Remove last character if it's a digit
df <- df %>%
 mutate(type = if_else(str_detect(str_sub(type, -1), "\\d"),
str_sub(type, 1, -2),type))



# Get the last two column names
last_two <- names(df)[(ncol(df)-1):ncol(df)]


# Get the remaining column names
remaining <- setdiff(names(df), last_two)

# Reorder: last two first, then the rest
df_reordered <- df %>%
 select(all_of(last_two), all_of(remaining))




SRestimates = df_reordered

rownames(SRestimates)= 1:nrow(SRestimates)


colnames(SRestimates)[4:6] =c("se","95_lower","95_upper")
SRestimates$type = trimws(SRestimates$type)
```


```{r}
# final$jackknife_1order


# Filter SRestimates for "1st order jackknife"
jackknife_1 <- SRestimates %>%
 filter(type == "1st order jackknife") %>%
 select(site_id, Estimate) %>%
 rename(jackknife_1order = Estimate)


jackknife_2 <- SRestimates %>%
 filter(type == "2nd order jackknife") %>%
 select(site_id, Estimate) %>%
 rename(jackknife_2order = Estimate)

chao_2 <- SRestimates %>%
 filter(type == "Chao2 (Chao, 1987)") %>%
 select(site_id, Estimate) %>%
 rename(chao_2 = Estimate)

chao_2_bc <- SRestimates %>%
 filter(type == "Chao2-bc") %>%
 select(site_id, Estimate) %>%
 rename(chao_2_bc = Estimate)

# Conditionally remove the column if it exists
cols_to_remove <- intersect(c("jackknife_1order", "jackknife_2order","chao_2","chao_2_bc"), names(final))

final <- final %>% select(-all_of(cols_to_remove))


# Join with final by site_id
final <- final %>%
 left_join(jackknife_1, by = "site_id") %>%
 left_join(jackknife_2, by = "site_id") %>%
 left_join(chao_2, by = "site_id") %>%
 left_join(chao_2_bc, by = "site_id")





final$jackknife_2order = SRestimates[SRestimates$type=="2nd order jackknife","Estimate"]
final$chao2   = SRestimates[SRestimates$type=="Chao2 (Chao, 1987)","Estimate"]
final$chao2bc = SRestimates[SRestimates$type=="Chao2-bc","Estimate"]
```





# Export results to excel
```{r}
# Create a blank workbook
OUT <- openxlsx::createWorkbook()

# Add some sheets to the workbook
openxlsx::addWorksheet(OUT, "SRestimates")
openxlsx::addWorksheet(OUT, "SRestimates_complete")

# Write the data to the sheets
openxlsx::writeData(OUT, sheet = "SRestimates", x = final)
openxlsx::writeData(OUT, sheet = "SRestimates_complete", x = SRestimates)


# Export the file
openxlsx::saveWorkbook(OUT, paste("results/",gsub("-",".",Sys.Date()),
                            "-SpeciesRarefactionCurves.xlsx",sep=""),
                       overwrite = TRUE)

```
