---
title: "RDA and Variation partitioning"
author: "Bernd Panassiti"
output: html_document
date: 'created: 23.04.2025, last modified: `r format(Sys.Date(), format="%d.%m.%Y")`'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r setup, include=FALSE}
library(knitr)
opts_knit$set(root.dir='../')                     # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center',                # aligns all figures
                echo=TRUE,                        # shows r-code
                message=FALSE,                    # suppresses library outputs
                warnings=FALSE,                   # suppresses library outputs
                tidy=TRUE,  # prevents the source code from running off a pdf page
                dev='pdf')
```

```{r}
#Install/load pacman
if(!require(pacman)){install.packages("pacman");require(pacman)}
#Install packages
p_load(tidyverse,dplyr,vegan,ggplot2,ggpubr,magrittr)
```

# load working data
```{r}
load("data/workingdata.RData")
```


# RDA
```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2025)

RDA <- rda(spec_RDA.sqrt, env_RDA[,-1], covar_RDA)
summary(RDA)

anova.cca(RDA, step = 1000) #globaler Test des RDA Ergebnisses
#-> das Ergebnis der RDA ist signifikant

ordiplot(RDA) #Ordiplot

smry <- summary(RDA)
RDAscoresSites  <- data.frame(smry$sites[,1:2])       # sites
RDAscoresSpecies  <- data.frame(smry$species[,1:2])     # species

```

1. RDA axis related to woody encroachment

# Variation Partioning
```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2025)



env_RDA.CovJuv <- env_RDA %>% 
  dplyr::select(Cover_Juvenile)

spe.part.all <- vegan::varpart(spec_RDA.sqrt, 
                        env_RDA.CovJuv,
                        covar_RDA)

plot(spe.part.all,
     Xnames = c( "Cover Juvenile", "Alm"), # name the partitions
     bg = c("seagreen3", "mediumpurple"), alpha = 80, # colour the circles
     digits = 2, # only show 2 digits
     cex = 1.5) #<- Venn diagramm mit Farben und Beschriftung
```


# Boxplots based on RDA scores

Waldartenliste
1.1: geschlossener Wald
1.2: Waldränder und –verlichtungen
2.1: Wald wie im Offenland
2.2: auch Wald, aber Schwerpunkt Offenland

B - Bäume, S - Sträucher, K - Kräuter, E - Epiphyten



```{r}
df = RDAscoresSites
df$plot_id = rownames(RDAscoresSites)

RDAscoresSitesTraits <- merge(df, traitsSummaryPlot, by = "plot_id", all.x = TRUE)
```




# Linear Regression: RDA1 vs Forage Value

```{r}
df = RDAscoresSitesTraits %>%
  mutate(forageValue = as.numeric(forageValue_weighted)) %>%      # convert FW to numeric
  rename(`Forage Value` = forageValue,
         `RDA1-Value` = RDA1) %>%      # rename for plotting
  filter(!is.na(`Forage Value`), !is.na(`RDA1-Value`))  # remove NA values


gRDA_forage = ggplot(df, aes(y = `Forage Value`, x = `RDA1-Value`)) +
  geom_point(color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "darkred", fill = "pink") +
  stat_regline_equation(
    aes(label = after_stat(eq.label)),
    label.x = -3,
    label.y = max(df$`Forage Value`, na.rm = TRUE)
  ) +
stat_cor(
  aes(label = after_stat(paste(rr.label, p.label, sep = "~`,`~"))),
  label.x = -3,
  label.y = max(df$`Forage Value`, na.rm = TRUE) - 0.2,
  p.accuracy = 0.001
)+
  theme_minimal() +
  labs(
    title = "",
    y = "Forage Value",
    x = "RDA1-Value"
  ); gRDA_forage
```





# Forest association
## Forest association - Frequencies
```{r}
df = RDAscoresSitesTraits %>%
  dplyr::select(RDA1,forestAssociation_1.1_freq,forestAssociation_2.1_freq)

df %<>%
  pivot_longer(
    cols = -RDA1,                     # 2. Keep 'species', pivot the rest
    names_to = "forestAssociation",                # 3. Column names become 'plot_id'
    values_to = "value"                  # 4. Values go into 'cover'
  )%>%
  mutate(forestAssociation = str_extract(forestAssociation, "\\d+\\.\\d+"))
  

df$forestAssociation = as.factor(df$forestAssociation)


# Plot
gRDA_forest  = ggplot(df, aes(y = value, x = RDA1,color=forestAssociation)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
    aes(label = after_stat(eq.label)),
    label.x = 1,
    label.y = c(4.3,4.6)
  ) +
stat_cor(
  aes(label = after_stat(paste(rr.label, p.label, sep = "~`,`~"))),
  label.x = 1,
  label.y = c(4.2,4.5 ),
  p.accuracy = 0.001
)+
  theme_minimal() +
  labs(
    title = "",
    y = "Forest association",
    x = "RDA1-Value"
  )+
  scale_color_manual("Treatments:",values = 
                       rev(ggsci::pal_jco("default")(3))[c(2,3)]); gRDA_forest
```



## Boxplot - Forest association

## Boxplot - Forest association
```{r}
df1 = RDAscoresSpecies
df1$species = rownames(RDAscoresSpecies)

traits_selected = traitsNew %>% 
  dplyr::filter(cover>0) %>%
  dplyr::select(plot_id,species,forestAssociation_1.1,forestAssociation_2.1,table33_34,table35)

                   

df2 <- traits_selected %>%
  pivot_longer(
    cols = -c("plot_id","species"),  # Alle Spalten außer "species"
    names_to = "type",
    values_to = "occ"
  )

df2 %<>% dplyr::filter(occ>0)

traitsForestAccBiotope_plotData <- merge(df1, df2, by = "species", all.x = TRUE)


df = traitsForestAccBiotope_plotData %>%  
   mutate(type = recode(type, "forestAssociation_1.1" = "1.1", "forestAssociation_2.1" = "2.1"))%>%
  dplyr::filter(type %in% c("1.1","2.1"))

df$type = as.factor(df$type)

df %<>% group_by(species,type)%>%
  dplyr::summarise(RDA1 = mean(RDA1))

# Calculate sample sizes
n_labels_wald <- df %>%
  group_by(type) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(y = min(df$RDA1, na.rm = TRUE) - 0.05)

# Plot
gRDA_forest = ggplot(df, aes(x = type, y = RDA1)) +
  geom_boxplot(outlier.shape = NA, fill = "white",
               color = "black") +
  stat_summary(fun = mean, geom = "point", 
               shape = 20, size = 3, color = "red") +
  stat_compare_means(method = "wilcox.test", 
                     label = "p.signif", 
                     comparisons = combn(levels(df$type), 2, 
                                         simplify = FALSE)) +
  geom_text(data = n_labels_wald,
            aes(x = type, y = y, label = paste("n =", n)),
            inherit.aes = FALSE, size = 3.5) +
  labs(x = "Forest association", y = "RDA1-Value", title = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));gRDA_forest
```






# Biotope
table 33 grasslands in intermediate to heigh elevations
table 34 grasslands in low to intermediate  elevations
table 35 dry grasslands


## Biotope - Frequencies
```{r}
df = RDAscoresSitesTraits %>%
  dplyr::select(RDA1,table33_34_freq,table35_freq)

df %<>%
  pivot_longer(
    cols = -RDA1,                     # 2. Keep 'species', pivot the rest
    names_to = "biotope",                # 3. Column names become 'plot_id'
    values_to = "value"                  # 4. Values go into 'cover'
  )%>%
  mutate(biotope = recode(biotope,
                            "table33_34_freq" = "tables 33 and 34",
                            "table35_freq" = "table 35"))
  

df$biotope = as.factor(df$biotope)




gRDA_biotope = ggplot(df, aes(y = value, x = RDA1,color=biotope)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
    aes(label = after_stat(eq.label)),
    label.x = 1,
    label.y = c(4.8, 5.1)
  ) +
stat_cor(
  aes(label = after_stat(paste(rr.label, p.label, sep = "~`,`~"))),
  label.x = 1,
  label.y = c(4.7, 5.0),
  p.accuracy = 0.001
)+
  theme_minimal() +
  labs(
    title = "",
    y = "Weighted biotope value",
    x = "RDA1-Value"
  )+
  scale_color_manual("Treatments:",values = 
                       rev(ggsci::pal_jco("default")(3))[c(2,3)]); gRDA_biotope
```





## Boxplot - Biotope
```{r}
df1 = RDAscoresSpecies
df1$species = rownames(RDAscoresSpecies)

traits_selected = traitsNew %>% 
  dplyr::filter(cover>0) %>%
  dplyr::select(plot_id,species,forestAssociation_1.1,forestAssociation_2.1,table33_34,table35)

                   

df2 <- traits_selected %>%
  pivot_longer(
    cols = -c("plot_id","species"),  # Alle Spalten außer "species"
    names_to = "type",
    values_to = "occ"
  )

df2 %<>% dplyr::filter(occ>0)

traitsForestAccBiotope_plotData <- merge(df1, df2, by = "species", all.x = TRUE)


df = traitsForestAccBiotope_plotData %>%  
  dplyr::filter(type %in% c("table33_34","table35"))

df$type = as.factor(df$type)

df %<>% group_by(species,type)%>%
  dplyr::summarise(RDA1 = mean(RDA1))

# Calculate sample sizes
n_labels_biotope <- df %>%
  group_by(type) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(y = min(df$RDA1, na.rm = TRUE) - 0.05)

# Plot
gRDA_biotope = ggplot(df, aes(x = type, y = RDA1)) +
  geom_boxplot(outlier.shape = NA, fill = "white",
               color = "black") +
  stat_summary(fun = mean, geom = "point", 
               shape = 20, size = 3, color = "red") +
  stat_compare_means(method = "wilcox.test", 
                     label = "p.signif", 
                     comparisons = combn(levels(df$type), 2, 
                                         simplify = FALSE)) +
  geom_text(data = n_labels_biotope,
            aes(x = type, y = y, label = paste("n =", n)),
            inherit.aes = FALSE, size = 3.5) +
  scale_x_discrete(labels = c("table33_34" = "Tables 33 and 34", "table35" = "Table 35"))+
  labs(x = "§30 biotope species", y = "RDA1-Value", title = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));gRDA_biotope
```



































# Table 33
Tafel 33: Alpine Rase
34 & 35 zusammengefasst: Grünland mäßig

frischer Standorte, Magerstandorte

Bewertungindex: bei 1, schon biotope als gesichert

hypothese: bei früher beweiundg verschiebt sich mehr richtung 34/35

```{r}
save(
  gRDA_forage,gRDA_biotope, gRDA_forest,
  file="results/RDA_plots.RData")
```

