---
title: "Descriptive analyses"
author: "Bernd Panassiti"
date: 'created: 28.11.2024, last modified: `r format(Sys.Date(), format="%d.%m.%Y")`'
output: pdf_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
opts_knit$set(root.dir='../')                     # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center',                # aligns all figures
                echo=TRUE,                        # shows r-code
                message=FALSE,                    # suppresses library outputs
                warnings=FALSE,                   # suppresses library outputs
                tidy=TRUE,  # prevents the source code from running off a pdf page
                dev='pdf')
```


```{r}
#Install/load pacman
if(!require(pacman)){install.packages("pacman");require(pacman)}
#Install packages
p_load(readxl,ggplot2,tidyr,dplyr,magrittr,
       DHARMa,
       glmmTMB,
       interactions,# interact_plot
       ggeffects,   # ggpredict
       ggrepel      # geom_text_repel
       )
```

# information
| Paket     | Überdispersion-Test | AIC   | Residuenanalyse | Negative Binomial |
| --------- | ------------------- | ---   | --------------- | ----------------- |
| `glmmPQL` | ❌ nicht zuverlässig | ❌  | eingeschränkt  |  ❌ (nur Poisson)   |
| `glmmTMB` | ✅                   | ✅   | ✅             | ✅                     |


Encroachment = original data
encroachment = scaled data


# load working data
```{r}
load("data/workingdata.RData")
```




# Encroachment as continuous variables
## data preparation

z-scores (standardized units)

```{r}
model_data = plot_data

          
model_data$elevation    <- as.numeric(scale(model_data$Elevation))
model_data$encroachment <- as.numeric(scale(model_data$Cover_Juvenile))
model_data$exposition   <- as.numeric(scale(model_data$Exposition))
model_data$slope        <- as.numeric(scale(model_data$Hangneigung))

model_data %<>% dplyr::rename(SR = Gesamtartenzahl, 
                              site_id=Alm,
                              Encroachment = Cover_Juvenile)   

model_data %<>% dplyr::select(site_id,SR,elevation,encroachment,exposition,slope,Encroachment)
```



## Poisson - Run model
```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2025)

full <- glmmTMB::glmmTMB(SR ~ elevation + slope + exposition + 
                                    encroachment + (1| site_id),
                                   family = poisson(), 
                                   data = model_data)

summary(full)
```



```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2025)

simres <- simulateResiduals(full)

# graphical diagnostic
plot(simres)

# Test for overdispersion
testDispersion(simres)
```


p > 0.05	no strong overdispersion ->	Model is OK with Poisson

# Model reduction


```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2025)

m1 <- glmmTMB(SR ~ elevation + slope + encroachment + (1 | site_id),
              family = poisson(), data = model_data)

m2 <- glmmTMB(SR ~ elevation + exposition + encroachment + (1 | site_id),
              family = poisson(), data = model_data)

m3 <- glmmTMB(SR ~ slope + exposition + encroachment + (1 | site_id),
              family = poisson(), data = model_data)

m4 <- glmmTMB(SR ~ slope + encroachment + (1 | site_id),
              family = poisson(), data = model_data)

m5 <- glmmTMB(SR ~ elevation + encroachment + (1 | site_id),
              family = poisson(), data = model_data)

m6 <- glmmTMB(SR ~ exposition + encroachment + (1 | site_id),
              family = poisson(), data = model_data)

m7 <- glmmTMB(SR ~ elevation + slope + (1 | site_id),
              family = poisson(), data = model_data)

m8 <- glmmTMB(SR ~ elevation + exposition + (1 | site_id),
              family = poisson(), data = model_data)

m9 <- glmmTMB(SR ~ slope + exposition + (1 | site_id),
              family = poisson(), data = model_data)
```

```{r}
AIC(full, m1, m2, m3, m4, m5, m6, m7, m8, m9)
```

> AIC(full, m1, m2, m3, m4, m5, m6, m7, m8, m9)
     df      AIC
full  6 939.1034
m1    5 937.7661
m2    5 944.5266
m3    5 937.4029
m4    4 936.1615
m5    4 943.0342
m6    4 943.1881
m7    4 940.8668
m8    4 948.5857
m9    4 940.7048

m1 has lowest AIC



m4 has the lowest AIC, which suggests it is the best model 
among the three in terms of the trade-off between fit and complexity.
The difference in AIC between models is small (less than 2),
which typically means the models are similarly supported by the data. 
However, m4 is slightly preferred.



# Check residuals of best models

## m4
```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2025)

simres <- simulateResiduals(m4)

# graphical diagnostic
plot(simres)

# Test for overdispersion
testDispersion(simres)
```


p > 0.05	no strong overdispersion ->	Model is OK with Poisson


Based on the residual diagnostic plot for model m4, here's what we can conclude:

The QQ plot shows that the residuals closely follow the expected distribution, suggesting no major issues with normality.
The residuals vs. predicted plot shows some quantile deviations, but the combined adjusted quantile tests are not significant (marked as "n.s.").

✅ Conclusion:
Yes, you can still use model m4. The quantile deviations observed are not statistically significant, so they do not invalidate the model. The diagnostics suggest that m4 is a reasonable fit for your data. 



## check interaction
```{r}
rm(.Random.seed, envir=globalenv())
set.seed(2025)

m4_interaction <- glmmTMB(SR ~ slope * encroachment + (1 | site_id),
              family = poisson(), data = model_data)
```


```{r}
AIC( m4, m4_interaction)
```


```{r}
anova(m4_interaction, m4, test = "Chisq")
```


The AIC is slightly higher for m4_interaction, which suggests no improvement in model fit despite the added complexity.
The Chi-squared test from anova() gives a p-value of 0.2402, which is not significant. This means the interaction term does not significantly improve the model.
✅ Conclusion:
m4 (without interaction) is the better model: it is simpler and not significantly worse in fit.
Stick with m4 unless you have a strong theoretical reason to include the interaction.


# Performance
## R²
```{r}
# Calculate R²
r2 <- performance::r2(m4)
print(r2)
```



An R² (R-squared) value of 0.07 means that your model explains 7% of the variance in the response variable — in your case, plant species richness (SR).

🔍 What this tells you:
7% of the variation in SR can be explained by the predictors in your model (e.g., slope and encroachment).
The remaining 93% of the variation is due to other factors not included in the model, random noise, or measurement error.

## Plot
```{r}
# Create new data for prediction
new_data <- data.frame(
      encroachment = seq(min(model_data$encroachment),
                         max(model_data$encroachment),
                         length.out = 100),
  slope = mean(model_data$slope),
  site_id = NA  # Random effect not used in fixed-effect prediction
)

# Predict with confidence intervals
pred <- predict(m4, newdata = new_data, type = "response", 
                se.fit = TRUE)
new_data$fit <- pred$fit
new_data$lower <- pred$fit - 1.96 * pred$se.fit
new_data$upper <- pred$fit + 1.96 * pred$se.fit


# back transform ecnroachment

mean_cover <- mean(model_data$Encroachment, na.rm = TRUE)
sd_cover <- sd(model_data$Encroachment, na.rm = TRUE)

new_data$Encroachment <- new_data$encroachment * sd_cover + mean_cover



# Extract coefficients
coefs <- fixef(m4)$cond
intercept <- round(coefs["(Intercept)"], 2)
enc_coef <- round(coefs["encroachment"], 2)

# R² and p-value
r2m <- round(performance::r2(m4)$R2_marginal, 2)
r2c <- floor(performance::r2(m4)$R2_conditional * 100) / 100
p_val <- format.pval(summary(m4)$coefficients$cond["encroachment", "Pr(>|z|)"], digits = 3, eps = 0.001)


# Create plain text annotation
annotation_text <- paste0(
  "y = ", intercept, " + ",  enc_coef, " × encroachment\n",
  "R² marginal = ", r2m, "\n",
  "R² conditional = ", r2c, "\n",
  "p = ", p_val
)


# Determine max values for positioning
x_pos <- max(model_data$Encroachment, na.rm = TRUE)
y_pos <- max(model_data$SR, na.rm = TRUE)

# Plot with right-aligned annotation
Fig2=ggplot(model_data, aes(x = Encroachment, y = SR)) +
  geom_point(alpha = 0.5) +
  geom_line(data = new_data, aes(x = Encroachment, y = fit), 
            color = "black",size = 1) +
  geom_ribbon(data = new_data, aes(x = Encroachment, 
                                   ymin = lower,
                                   ymax = upper),
              fill = "grey", alpha = 0.3, inherit.aes = FALSE) +
  annotate("text", x = x_pos, y = y_pos, label = annotation_text,
           hjust = 1, vjust = 1, size = 4) +
  labs(x = "Encroachment (%)", y = "Plant species richness",
       title = "") +
  theme_minimal(); Fig2
```


# Final regession table
```{r}
# Extract fixed effects
m4_summary <- broom.mixed::tidy(m4, effects = "fixed")

# Round and format for Excel output
m4_summary$estimate <- round(m4_summary$estimate, 3)
m4_summary$std.error <- round(m4_summary$std.error, 3)
m4_summary$statistic <- round(m4_summary$statistic, 2)
m4_summary$p.value <- signif(m4_summary$p.value, 3)

# Optional: add significance stars
m4_summary$significance <- cut(m4_summary$p.value,
                                   breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
                                   labels = c("***", "**", "*", ".", ""))


openxlsx::write.xlsx(m4_summary, file = 
                                paste("results/",gsub("-",".",
                                    Sys.Date()),
                                    "-GLMM_summary_table.xlsx",sep="")
                     , rowNames = FALSE)
```



# save
```{r}
save(
  model_data,
  m4,m4_summary,
  Fig2,
  file="results/regression_results.RData"
)
```

